AEC ANALYSIS R SCRIPT
=====================
Author: Steve McEachern
Date: 

```{r }
date()
```

github: git@github.com:stevenmce/aec_analysis.git
This is a special R script which can be used to generate a report. You can
write normal text in roxygen comments.

First we set up some options (you do not have to do this):

```{r setup, include=FALSE}
require(knitr)
opts_chunk$set(fig.path='figure/AEC-')

# Enable required libraries
require(Hmisc)
require(foreign)
require(XML)

```{r startReport}
```

The report begins here.

```{r }
# This code reads in the data

```{r readData, cache=TRUE}
```

READ IN DATA
============
This section reads in the data files
FIRST PREFERENCE VOTES 2010 BY _Candidate_ CSV file
===================================================
Code from Simon Jackman - source is comments section of:
http://jackman.stanford.edu/blog/?p=1760 
LOAD
----
Define the location of the data (AEC website)

```{r }
theURL <- "http://results.aec.gov.au/15508/Website/Downloads/HouseFirstPrefsByCandidateByVoteTypeDownload-15508.csv"
```

Read in the data

```{r }
First_Preferences_By_Candidate_By_Vote_Type <- read.csv(url(theURL),
                     skip=1,
                     header=TRUE,
                     quote="\"",
                     stringsAsFactors=FALSE)
```

CLEAN
-----
Vote counts occuring by type
Define variables

```{r }
vars <- c("OrdinaryVotes","AbsentVotes","ProvisionalVotes",
          "PrePollVotes","PostalVotes","TotalVotes")
```

DO
--
Aggregate votes by division (sum all candidates)

```{r }
out <- NULL
for(v in vars){
  theVar <- First_Preferences_By_Candidate_By_Vote_Type[,match(v,names(First_Preferences_By_Candidate_By_Vote_Type))]
  out <- cbind(out,
               tapply(theVar,First_Preferences_By_Candidate_By_Vote_Type$DivisionNm,sum,na.rm=TRUE))
}
```

Define variable names

```{r }
outvars <-  c("OrdinaryVotes","AbsentVotes","ProvisionalVotes",
              "PrePollVotes","PostalVotes","TotalVotes","DivisionName")

z <- apply(out,2,function(x)sum(x!=0))
names(z) <- vars
print(z)

## Swing against ALP candidates 
## THIS CODE TO GENERATE SWING DOESNT WORK??
#swing <- aggregate(First_Preferences_By_Candidate_By_Vote_Type, list(Party = First_Preferences_By_Candidate_By_Vote_Type["PartyAb" = "ALP"]), mean)
#print(swing)
```

Convert out data matrix to a data frame

```{r }
outdf <- as.data.frame(out)
```

Transform division from row.names to a variable 
http://stackoverflow.com/questions/11427434/how-to-create-a-variable-of-rownames

```{r }
outdf$Division = rownames(outdf)
rownames(outdf) = NULL
```

Assign column (variable) names

```{r }
colnames(outdf) <- outvars
```

Describe the data

```{r }
describe(outdf)
```

CLEAN UP
--------
Rename the data frame

```{r }
jackman_votetype_by_division <- outdf
remove(outdf)
remove(out)
```

FILE TWO
========
FIRST PREFERENCE RESULTS BY _Division_ - HTML table
IF REQUIRED:
remove(HouseFirstPrefsTppByDivision)

LOAD
----
READ IN DATA - NOTE StringsAsFactors - False to avoid FACTOR problems

```{r }
HouseFirstPrefsTppByDivision <- readHTMLTable("http://results.aec.gov.au/15508/Website/HouseFirstPrefsTppByDivision-15508-NAT.htm",
                            header = c("Division","State","ALP_1Prf","LP_1Pref","LNQ_1Pref","GRN_1Pref","NP_1Pref","OTH_1Pref","ALP_2PP","LNP_2PP"),
                            skip.rows=c(1,2), trim=TRUE, as.data.frame=TRUE, which=5, stringsAsFactors=FALSE)
```

Describe the data

```{r }
describe(HouseFirstPrefsTppByDivision)
```

Apply variable names to columns

```{r }
HouseFirstPrefsTppByDivision_varnames <-  c("Division","State","ALP_1Prf","LP_1Pref","LNQ_1Pref","GRN_1Pref","NP_1Pref","OTH_1Pref","ALP_2PP","LNP_2PP")
colnames(HouseFirstPrefsTppByDivision) <- HouseFirstPrefsTppByDivision_varnames
```

CLEAN
-----
Recode values of NATIONAL TOTAL due to missing State value
select rows where v1 is 99 and recode column v1 

```{r }
HouseFirstPrefsTppByDivision[HouseFirstPrefsTppByDivision$Division=="National Total","State"] <- "National"
HouseFirstPrefsTppByDivision[HouseFirstPrefsTppByDivision$Division=="National Total","ALP_1Prf"] <- 37.99
HouseFirstPrefsTppByDivision[HouseFirstPrefsTppByDivision$Division=="National Total","LP_1Pref"] <- 30.46
HouseFirstPrefsTppByDivision[HouseFirstPrefsTppByDivision$Division=="National Total","LNQ_1Pref"] <- 9.12
HouseFirstPrefsTppByDivision[HouseFirstPrefsTppByDivision$Division=="National Total","GRN_1Pref"] <- 11.76
HouseFirstPrefsTppByDivision[HouseFirstPrefsTppByDivision$Division=="National Total","NP_1Pref"] <- 3.73
HouseFirstPrefsTppByDivision[HouseFirstPrefsTppByDivision$Division=="National Total","OTH_1Pref"] <- 6.94
HouseFirstPrefsTppByDivision[HouseFirstPrefsTppByDivision$Division=="National Total","ALP_2PP"] <- 50.12
HouseFirstPrefsTppByDivision[HouseFirstPrefsTppByDivision$Division=="National Total","LNP_2PP"] <- 49.88

#CHECK CHANGES APPLIED CORRECTLY
#VALUES: National Total [STATE]  37.99  30.46  9.12  11.76  3.73  6.94  50.12  49.88
tail(HouseFirstPrefsTppByDivision)

#CREATE A State FACTOR VARIABLE
HouseFirstPrefsTppByDivision$State_factor <- as.factor(HouseFirstPrefsTppByDivision$State)

#RENAME THE DIVISION VARIABLE (FOR FILE MERGING)
HouseFirstPrefsTppByDivision$DivisionNm <- as.character(HouseFirstPrefsTppByDivision$Division)
HouseFirstPrefsTppByDivision$Division <- NULL
```

FILE THREE - HOUSE VOTES COUNTED BY DIVISION
============================================
LOAD
----
Define the location of the data (AEC website)

```{r }
theURL <- "http://results.aec.gov.au/15508/Website/Downloads/HouseVotesCountedByDivisionDownload-15508.csv"

## Read in the data
HouseVotesCountedByDivision <- read.csv(url(theURL),
                               skip=1,
                               header=TRUE,
                               quote="\"",
                               stringsAsFactors=FALSE)

require(ggplot2)
quickplot(HouseVotesCountedByDivision$TotalVotes)

# Run the following on the command line to generate R-markdown document
# spin(aec_read_data_knitr.r)
```

END OF FILE
