AEC Data Analysis
========================================================

This is my first attempt at an R Markdown document. 

When you click the **Knit HTML** button a web page will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

## FUNC??

This code loads the relevant packages into R.

```{r loadLibraries}
## Enable required libraries
library(Hmisc)
library(foreign)
library(XML)
```

## LOAD

This code reads in the first set of data.

```{r readData}
## READ IN DATA

## FIRST PREFERENCE VOTES 2010 BY _Candidate_ CSV file
## Code from Simon Jackman - source is comments section of:
## http://jackman.stanford.edu/blog/?p=1760 

## Define the location of the data (AEC website)
theURL <- "http://results.aec.gov.au/15508/Website/Downloads/HouseFirstPrefsByCandidateByVoteTypeDownload-15508.csv"

## Read in the data
First_Preferences_By_Candidate_By_Vote_Type <- read.csv(url(theURL),
                     skip=1,
                     header=TRUE,
                     quote="\"",
                     stringsAsFactors=FALSE)

## READ IN DATA

## FIRST PREFERENCE VOTES 2010 BY _Candidate_ CSV file
## Code from Simon Jackman - source is comments section of:
## http://jackman.stanford.edu/blog/?p=1760 

## Define the location of the data (AEC website)
theURL <- "http://results.aec.gov.au/15508/Website/Downloads/HouseFirstPrefsByCandidateByVoteTypeDownload-15508.csv"

## Vote counts occuring by type
vars <- c("OrdinaryVotes","AbsentVotes","ProvisionalVotes",
          "PrePollVotes","PostalVotes","TotalVotes")

out <- NULL
for(v in vars){
  theVar <- First_Preferences_By_Candidate_By_Vote_Type[,match(v,names(First_Preferences_By_Candidate_By_Vote_Type))]
  out <- cbind(out,
               tapply(theVar,First_Preferences_By_Candidate_By_Vote_Type$DivisionNm,sum,na.rm=TRUE))
}

outvars <-  c("OrdinaryVotes","AbsentVotes","ProvisionalVotes",
              "PrePollVotes","PostalVotes","TotalVotes","DivisionName")

z <- apply(out,2,function(x)sum(x!=0))
names(z) <- vars
print(z)

## Swing against ALP candidates 
## THIS CODE TO GENERATE SWING DOESNT WORK??
#swing <- aggregate(First_Preferences_By_Candidate_By_Vote_Type, list(Party = First_Preferences_By_Candidate_By_Vote_Type["PartyAb" = "ALP"]), mean)
#print(swing)

# Convert out data matrix to a data frame
outdf <- as.data.frame(out)

# Transform division from row.names to a variable 
# http://stackoverflow.com/questions/11427434/how-to-create-a-variable-of-rownames
outdf$Division = rownames(outdf)
rownames(outdf) = NULL

# Assign column (variable) names
colnames(outdf) <- outvars

# Describe the data
describe(outdf)

# Rename the data frame
jackman_votetype_by_division <- outdf
remove(outdf)
remove(out)

```

## DO

### Vote Distribution

Plot the vote distribution by division

```{r plotTest}
library(ggplot2)
qplot(jackman_votetype_by_division$TotalVotes,
      main="Total Votes by Division, 2010 Federal Election",
      xlab="Total votes")

```

## END

End of my first file

